---
title: "Y"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The question you are answering

You use the `##` to create a header. Use the header to indicate the question you are answering. After the header, in this space here, annotate your answer, telling us what you did.

```{r chunk name}
# Here is where your R code goes. ``` opens the code chunk, followed by {r}. Then ``` closes the chunk. When you 'knit' the file, the codes will evaluate.

gapminder <- read.csv(here::here("data/gapminder5.csv"), stringsAsFactors=FALSE)
```

Use an if() statement to print a suitable message reporting whether there are any records from 2002 in the gapminder dataset. Now do the same for 2012.

Hint: use the any function.

```{r another chunk name}
year<-2002
if(any(gapminder$year == year)){
   print(paste("Record(s) for the year",year,"found."))
} else {
  print(paste("No records for year",year))
}

year1<-2012
if(any(gapminder$year == year1)){
   print(paste("Record(s) for the year",year1,"found."))
} else {
  print(paste("No records for year",year1))
}
```
Write a script that finds the mean life expectancy by country for countries whose population is below the mean for the dataset

```{r}
overall_pop <- mean(gapminder$pop)
for (i in unique(gapminder$country)) {
   country_pop_mean <- mean(gapminder$pop[gapminder$country==i])
   if (country_pop_mean < overall_pop) {
    country_exp_mean <- mean(gapminder$lifeExp[gapminder$country==i])
     print(paste("Mean Life Expectancy of", i, "is", country_exp_mean))
   } 
} 
```
Write a script that loops through the gapminder data by continent and prints out whether the mean life expectancy is smaller than 50, between 50 and 70, or greater than 70.
```{r}
threshold_1<-50
threshold_2<-70
for (i in unique(gapminder$continent)) {
    mean_life_exp<-mean(gapminder$lifeExp[gapminder$continent==i])
    if (mean_life_exp<threshold_1) {
        print (paste("continent", i, "has mean life expectancy smaller than 50"))
    }
    if (threshold_1<=mean_life_exp & mean_life_exp<=threshold_2) {
        print(paste("continent", i, "has mean life expectancy between 50 and 70"))
    }
    if (mean_life_exp>threshold_2){
        print(paste("continent", i, "has mean life expectancy larger than 70"))
    }
}

```
Create a function that given a data frame will print the name of each column and the class of data it contains. Use the gapminder dataset. Hint: Use mode() or class() to get the class of the data in each column. Remember that names() or colnames() returns the name of the columns in a dataset.
```{r}
name_class <- function(a){
    col_names<-names(a)
    for (i in col_names) {
        result<-class(a[[i]])
        print(paste(i, ":", result))
    }
}

name_class(gapminder)
```

Create a function that given a vector will print the mean and the standard deviation of a vector, it will optionally also print the median. Hint: include an argument that takes a boolean (TRUE/FALSE) operator and then include an if statement.
```{r}
mean_std <- function(a, med=F){
    print(mean(a))
    print(sd(a))
    if (med==T){
        print (median(a))
    }
}

vec<-c(1:5)
mean_std(vec, med=T)
mean_std(vec, med=F)
```
Use what you???ve learned so far to answer the following questions using the gapminder dataset. Be sure to include some visualizations!

What is the relationship between GDP per capita and life expectancy? Does this relationship change over time? (Hint: Use the natural log of both variables.)
```{r}
x<-log(gapminder$gdpPercap)
y<-log(gapminder$lifeExp)
plot(x,y)

```
The natural logs of two variables show a positive relationship over time. 

Does the relationship between GDP per capita and life expectacy vary by continent? Make sure you divide the Americas into North and South America.
```{r}
for (i in unique(continent)){
    life_exp_cont<- gapminder$lifeExp[gapminder$continent==i]
    GDP_cont<- gapminder$gdpPercap[gapminder$continent==i]
    plot(GDP_cont,life_exp_cont)
}

```
